---
title: "plotIGCSim"
author: "Xiang Ji"
date: "February 18, 2016"
output: html_document
---

This R script is used for generating plots, tables, summaries for writing the IGC draft.

```{r}
rm(list=ls())  # clean up workspace
# Now load all data
source("/Users/xji3/GitFolders/IGCSimulation/ReadData.R")
results.path <- "/Users/xji3/GitFolders/IGCSimulation/"
library(ggplot2)
```

__Branch length comparison__

Compare all post-duplication branches of therir IGC estimates, PAML estimates from the true value used in simulation

```{r}
# IGC Branch
for (branch.iter in 1:length(IGC.branch.list)){
  target <- NULL
  for (IGC.geo in IGC.geo.list){
    target <- rbind(target, c(true.blen[true.branch.list[branch.iter]], 
                              mean(get(paste("geo_", toString(IGC.geo), ".0_summary", sep = "" ))[IGC.branch.list[branch.iter], ]),
                              (sd(get(paste("geo_", toString(IGC.geo), ".0_summary", sep = "" ))[IGC.branch.list[branch.iter], ]) 
                               / sqrt(length(get(paste("geo_", toString(IGC.geo), ".0_summary", sep = "" ))[IGC.branch.list[branch.iter], ]))), 
                              (sd(get(paste("geo_", toString(IGC.geo), ".0_summary", sep = "" ))[IGC.branch.list[branch.iter], ]) 
                               )))
    }
  target <- data.frame(target)
  target <- cbind(IGC.assign.branch.list[branch.iter], "IGC", 
                  IGC.geo.list, target)
  colnames(target) <- c( "Branch", "Method", "IGC.geo", "True_blen", "estimated_blen", "se_of_blen", "sd_of_blen")
  assign(paste(IGC.assign.branch.list[branch.iter], "blen", sep = "."), target)
  }

# paralog1 Branch
for (branch.iter in 1:length(IGC.branch.list)){
  target <- NULL
  for (IGC.geo in IGC.geo.list){
    target <- rbind(target, c(true.blen[true.branch.list[branch.iter]], 
                              mean(get(paste("geo_", toString(IGC.geo), ".0_PAML_maxlnL_summary", sep = "" ))[PAML.paralog1.branch.list[branch.iter], ]),
                              (sd(get(paste("geo_", toString(IGC.geo), ".0_PAML_maxlnL_summary", sep = "" ))[PAML.paralog1.branch.list[branch.iter], ]) 
                               / sqrt(length(get(paste("geo_", toString(IGC.geo), ".0_PAML_maxlnL_summary", sep = "" ))[PAML.paralog1.branch.list[branch.iter], ]))), 
                              (sd(get(paste("geo_", toString(IGC.geo), ".0_PAML_maxlnL_summary", sep = "" ))[PAML.paralog1.branch.list[branch.iter], ]) 
                               )))
    }
  target <- data.frame(target)
  target <- cbind(IGC.assign.branch.list[branch.iter], "Paralog 1", IGC.geo.list, target)
  colnames(target) <- c("Branch", "Method", "IGC.geo", "True_blen", "estimated_blen", "se_of_blen", "sd_of_blen")
  assign(paste(IGC.assign.branch.list[branch.iter], "blen", sep = "."), rbind(get(paste(IGC.assign.branch.list[branch.iter], "blen", sep = ".")), target))
  }

# paralog2 Branch
for (branch.iter in 1:length(IGC.branch.list)){
  target <- NULL
  for (IGC.geo in IGC.geo.list){
    target <- rbind(target, c(true.blen[true.branch.list[branch.iter]], 
                              mean(get(paste("geo_", toString(IGC.geo), ".0_PAML_maxlnL_summary", sep = "" ))[PAML.paralog2.branch.list[branch.iter], ]),
                              (sd(get(paste("geo_", toString(IGC.geo), ".0_PAML_maxlnL_summary", sep = "" ))[PAML.paralog2.branch.list[branch.iter], ]) 
                               / sqrt(length(get(paste("geo_", toString(IGC.geo), ".0_PAML_maxlnL_summary", sep = "" ))[PAML.paralog2.branch.list[branch.iter], ]))), 
                              (sd(get(paste("geo_", toString(IGC.geo), ".0_PAML_maxlnL_summary", sep = "" ))[PAML.paralog2.branch.list[branch.iter], ]) 
                               )))
    }
  target <- data.frame(target)
  target <- cbind(IGC.assign.branch.list[branch.iter], "Paralog 2", IGC.geo.list, target)
  colnames(target) <- c("Branch", "Method", "IGC.geo", "True_blen", "estimated_blen", "se_of_blen", "sd_of_blen")
  assign(paste(IGC.assign.branch.list[branch.iter], "blen", sep = "."), rbind(get(paste(IGC.assign.branch.list[branch.iter], "blen", sep = ".")), target))
  }

for (IGC.geo in IGC.geo.list){
  target <- NULL
  for (branch.iter in 1:length(IGC.branch.list)){
    branch.target <- get(paste(IGC.assign.branch.list[branch.iter], "blen", sep = "."))
    target <- rbind(target, branch.target[branch.target[, "IGC.geo"] == IGC.geo, ])
    }
  target <- data.frame(target)
  assign(paste("geo_", toString(IGC.geo), ".0_blen_summary", sep = "" ), target)
  }

```


Now plot estimated blen vs. true blen of each branches, change error bar to reflect 95% credit interval

```{r}
for (IGC.geo in IGC.geo.list){
  
  #http://gis.stackexchange.com/questions/120623/inserting-a-zoomed-in-plot-with-a-full-scale-plot-in-the-sp-package-in-r 
  # settings for full plot
  fulltheme <- theme(#panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
    #panel.background = element_blank() 
    #axis.text.x=element_blank(), axis.text.y=element_blank(),
    #axis.ticks=element_blank(),
    #axis.title.x=element_blank(), axis.title.y=element_blank()
    )
  
  # settings for zoom plot
  zoomtheme <- theme(legend.position="none", 
                     #axis.line=element_blank(),axis.text.x=element_blank(),
                     #axis.text.y=element_blank(),axis.ticks=element_blank(),
                     axis.title.x=element_blank(),axis.title.y=element_blank(),
                     panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                     #plot.margin = unit(c(0,0,-0.6,-0.6),"mm"),                   
                     panel.background = element_rect(color='red', fill="white")
                     )
  
  target <- get(paste("geo_", toString(IGC.geo), ".0_blen_summary", sep = "" ))
  p.full <- ggplot(target, aes(x = True_blen, y= estimated_blen, colour = Method, group = Method)) +
    geom_errorbar(aes(ymin = estimated_blen -  qnorm(0.975) * se_of_blen, ymax = estimated_blen + qnorm(0.975) * se_of_blen)) + 
    #    geom_line() + 
    geom_point(shape = 21, fill = "white") + 
    geom_abline(intercept = 0.0, slope = 1.0, lty = 2, colour = "grey") +
    xlab("True Branch Length") + 
    ylab("Estimated Branch Length") + 
    xlim(0.0, 0.25) + #max(target[, "True_blen"], target[, "estimated_blen"] + target[, "se_of_blen"]) * 1.1) + 
    ylim(0.0, 0.25) + #max(target[, "True_blen"], target[, "estimated_blen"] + target[, "se_of_blen"]) * 1.1) +
    ggtitle(paste("estimated blen v.s. true blen tract length = ", toString(IGC.geo), " NOT respect codon boundary", sep = "")) +
    fulltheme
  
  p.zoom <- ggplot(target, aes(x = True_blen, y= estimated_blen, colour = Method, group = Method)) +
    geom_errorbar(aes(ymin = estimated_blen - qnorm(0.975) * se_of_blen, ymax = estimated_blen + qnorm(0.975) * se_of_blen)) + 
    #    geom_line() + 
    geom_point(shape = 21, fill = "white") + 
    geom_abline(intercept = 0.0, slope = 1.0, lty = 2, colour = "grey") +
    coord_cartesian(xlim = c(0.0, 0.05), ylim = c(0.0, 0.05)) +
    xlab("True Branch Length") + 
    ylab("Estimated Branch Length") + 
    zoomtheme
  g <- ggplotGrob(p.zoom)
  
  x.off <- 0.15
  plot.together <- p.full + annotation_custom(grob = g, xmin = 0.125, xmax = 0.25, 
                                              ymin = 0.0, 
                                              ymax = 0.125) 
  
  ggsave(paste(results.path, "estimated blen vs true blen geo_", toString(IGC.geo), ".pdf", sep = ""))
  ggsave(paste(results.path, "estimated blen vs true blen geo_", toString(IGC.geo), ".jpg", sep = ""))
  
  }

```

__Tau estimate comparison__

```{r}
Tau.summary <- NULL
for (IGC.geo in IGC.geo.list){
  target <- get(paste("geo_", toString(IGC.geo), ".0_summary", sep = ""))
  Tau.summary <- rbind(Tau.summary, c(IGC.geo, mean(target["tau", ]), 
                                      sd(target["tau", ]) / sqrt(dim(target)[2]), 
                                      sd(target["tau", ])))
  }
colnames(Tau.summary) <- c("IGC.geo", "mean", "se", "sd")
Tau.summary <- data.frame(Tau.summary)
ggplot(Tau.summary, aes(x = IGC.geo, y= mean)) +
  geom_errorbar(aes(ymin = mean - qnorm(0.975) * sd, ymax = mean + qnorm(0.975) * sd)) + 
  #  geom_line() + 
  geom_point(shape = 21, fill = "white") + 
  xlab("Expected tract length") + 
  ylab("Estimated tau") + 
  #    xlim(0.0, max(target[, "True_blen"], target[, "estimated_blen"] + target[, "se_of_blen"]) * 1.1) + 
  #    ylim(0.0, max(target[, "True_blen"], target[, "estimated_blen"] + target[, "se_of_blen"]) * 1.1) +
  geom_hline(yintercept = 1.409408, lty = 2) + 
  ggtitle("estimated Tau value NOT respect codon boundary")
ggsave(paste(results.path, "estimated tau NOT respect boundary summary.pdf", sep = ""))

```

__Kappa estimate comparison__

```{r}
Kappa.summary <- NULL
for (IGC.geo in IGC.geo.list){
  target <- get(paste("geo_", toString(IGC.geo), ".0_summary", sep = ""))
  Kappa.summary <- rbind(Kappa.summary, c(IGC.geo, mean(target["kappa", ]), sd(target["kappa", ]) / sqrt(dim(target)[2])))
  }
colnames(Kappa.summary) <- c("IGC.geo", "mean", "se")
Kappa.summary <- data.frame(Kappa.summary)
ggplot(Kappa.summary, aes(x = IGC.geo, y= mean)) +
  geom_errorbar(aes(ymin = mean - qnorm(0.975) * se, ymax = mean + qnorm(0.975) * se)) + 
  #  geom_line() + 
  geom_point(shape = 21, fill = "white") + 
  xlab("Mean tract length") + 
  ylab("Estimated Kappa") + 
  #    xlim(0.0, max(target[, "True_blen"], target[, "estimated_blen"] + target[, "se_of_blen"]) * 1.1) + 
  #    ylim(0.0, max(target[, "True_blen"], target[, "estimated_blen"] + target[, "se_of_blen"]) * 1.1) +
  geom_hline(yintercept = 8.4043336, lty = 2) + 
  ggtitle("estimated Kappa value NOT respect codon boundary")
ggsave(paste(results.path, "estimated kappa NOT respect boundary summary.pdf", sep = ""))


```


__omega estimate comparison__

```{r}
omega.summary <- NULL
for (IGC.geo in IGC.geo.list){
  target <- get(paste("geo_", toString(IGC.geo), ".0_summary", sep = ""))
  omega.summary <- rbind(omega.summary, c(IGC.geo, mean(target["omega", ]), sd(target["omega", ]) / sqrt(dim(target)[2])))
  }
colnames(omega.summary) <- c("IGC.geo", "mean", "se")
omega.summary <- data.frame(omega.summary)
ggplot(omega.summary, aes(x = IGC.geo, y= mean)) +
  geom_errorbar(aes(ymin = mean - qnorm(0.975) * se, ymax = mean + qnorm(0.975) * se)) + 
  #  geom_line() + 
  geom_point(shape = 21, fill = "white") + 
  xlab("Mean tract length") + 
  ylab("Estimated omega") + 
  #    xlim(0.0, max(target[, "True_blen"], target[, "estimated_blen"] + target[, "se_of_blen"]) * 1.1) + 
  #    ylim(0.0, max(target[, "True_blen"], target[, "estimated_blen"] + target[, "se_of_blen"]) * 1.1) +
  geom_hline(yintercept = 1.0, lty = 2) + 
  ggtitle("estimated omega value NOT respect codon boundary")
ggsave(paste(results.path, "estimated omega NOT respect boundary summary.pdf", sep = ""))
```

Mar 4th 2016 update
Now plot tau, kappa, omega in the same plot

```{r}
tau.kappa.omega.summary <- rbind(cbind(omega.summary, "variable" = "omega"), 
                                 cbind(Kappa.summary, "variable" = "kappa"),
                                 cbind(Tau.summary,   "variable" = "tau"))
tau.kappa.omega.summary <- data.frame(tau.kappa.omega.summary)
ggplot(tau.kappa.omega.summary, aes(x = IGC.geo, y= mean, colour = variable, group = variable)) +
  geom_errorbar(aes(ymin = mean - qnorm(0.975) * se, ymax = mean + qnorm(0.975) * se)) + 
  #  geom_line() + 
  geom_point(shape = 21, fill = "white") + 
  xlab("Mean tract length") + 
  ylab("Estimate") + 
  ylim(0.0, 10.0) +
  geom_hline(yintercept = c(1.0, 8.4043336, 1.409408), lty = 2, colour = c("red", "green", "blue")) + 
  ggtitle("Variable estimated values NOT respect codon boundary")
ggsave(paste(results.path, "estimated variables NOT respect boundary summary.pdf", sep = ""))


```


March 11th 2016 update

Add proportion of changes due to IGC plots

```{r}
percent.summary <- NULL
for (IGC.geo in IGC.geo.list){
  target <- get(paste("percent.IGC.geo.", toString(IGC.geo), ".0", sep = ""))
  percent.summary <- rbind(percent.summary, c(IGC.geo, mean(target), sd(target)))
  }
colnames(percent.summary) <- c("IGC.geo", "mean", "sd")
percent.summary <- data.frame(percent.summary)

ggplot(percent.summary, aes(x = IGC.geo, y= mean)) +
  geom_errorbar(aes(ymin = mean - qnorm(0.975) * sd, ymax = mean + qnorm(0.975) * sd)) + 
  #  geom_line() + 
  geom_point(shape = 21, fill = "white") + 
  xlab("Expected tract length") + 
  ylab("Estimated proportion of changes due to IGC") + 
  #    xlim(0.0, max(target[, "True_blen"], target[, "estimated_blen"] + target[, "se_of_blen"]) * 1.1) + 
  #    ylim(0.0, max(target[, "True_blen"], target[, "estimated_blen"] + target[, "se_of_blen"]) * 1.1) +
  geom_hline(yintercept = 0.2131, lty = 2) + 
  ggtitle("estimated proportion of changes due to IGC NOT respect codon boundary")
ggsave(paste(results.path, "estimated proportion of changes due to IGC NOT respect boundary summary.pdf", sep = ""))
ggsave(paste(results.path, "estimated proportion of changes due to IGC NOT respect boundary summary.jpg", sep = ""))

```

Now re-visit N0_N1 branch

```{r}
 p.full <- ggplot(N0.N1.blen, aes(x = IGC.geo, y= estimated_blen, colour = Method, group = Method)) +
    geom_errorbar(aes(ymin = estimated_blen -  qnorm(0.975) * sd_of_blen, ymax = estimated_blen + qnorm(0.975) * sd_of_blen)) + 
    #    geom_line() + 
    geom_point(shape = 21, fill = "white") + 
    geom_hline(yintercept = 0.01972409, lty = 2) + 
    xlab("Mean tract length") + 
    ylab("Estimated Branch Length") + 
    #xlim(0.0, 0.25) + #max(target[, "True_blen"], target[, "estimated_blen"] + target[, "sd_of_blen"]) * 1.1) + 
    #ylim(0.0, 0.25) + #max(target[, "True_blen"], target[, "estimated_blen"] + target[, "sd_of_blen"]) * 1.1) +
    ggtitle(paste("estimated blen vs true blen N0_N1 NOT respect codon boundary", sep = "")) +
    fulltheme
ggsave(paste(results.path, "estimated blen vs true blen N0_N1.pdf", sep = ""))
ggsave(paste(results.path, "estimated blen vs true blen N0_N1.jpg", sep = ""))
```

Now calculate correlation of N0_N1 branch estimate between 2 different methods

```{r}
correlation.summary <- NULL
for (IGC.geo in IGC.geo.list){
  target.1 <- get(paste("geo_", toString(IGC.geo), ".0_summary", sep = ""))
  target.2 <- get(paste("geo_", toString(IGC.geo), ".0_PAML_maxlnL_summary", sep = ""))
  correlation.summary <- rbind(correlation.summary, c(IGC.geo, cor(target.1["(N0,N1)", ], target.2["N0_N1",]), 
                                              cor(target.1["(N0,N1)", ], target.2["N0_N6",])))
  }
colnames(correlation.summary) <- c("IGC.geo", "correlation with paralog 1", "correlation with paralog 2")


```

Now perform t-test of N0_N1 brach from PAML

```{r}
t.test.summary <- NULL
for (IGC.geo in IGC.geo.list){
  target.1 <- get(paste("geo_", toString(IGC.geo), ".0_summary", sep = ""))
  target.2 <- get(paste("geo_", toString(IGC.geo), ".0_PAML_maxlnL_summary", sep = ""))
  t.test.IGC <- t.test(target.1["(N0,N1)", ] - 0.01972409)
  t.test.paralog.1 <- t.test(target.2["N0_N1",] - 0.01972409)
  t.test.paralog.2 <- t.test(target.2["N0_N6",] - 0.01972409)
  t.test.summary <- rbind(t.test.summary, 
                               c(IGC.geo, t.test.IGC$statistic[[1]], t.test.paralog.1$statistic[[1]], 
                                 t.test.paralog.2$statistic[[1]], t.test.IGC$p.value, 
                                 t.test.paralog.1$p.value, t.test.paralog.2$p.value))
  
}
colnames(t.test.summary) <- c("IGC.geo", "t of IGC", "t of paralog 1", "t of paralog 2",
                              "p-value of IGC", "p-value of paralog 1", "p-value of paralog 2")

```

Now check if the under-estimation of N0_N1 is consistent with the over-estimation of outgroup branch

```{r}
# if it's consistent then, the estimate of outgroup branch from PAML should be higher 
# than that in IGC+MG94 model or true value by 1/2*sf
# similarly, the estimate of N0_N1 (and N0_N6 branch of paralog 2) should be less than
# that in IGC+MG94 model or true value by 1/2*sf
# true branch length of N0_kluyveri used in simulation is 0.215682181791
# true branch length of N0_N1       used in simulation is 0.01972409
N0.kluyveri.PAML.summary <- NULL
for (IGC.geo in IGC.geo.list){
  target <- get(paste("geo_", toString(IGC.geo), ".0_PAML_maxlnL_summary", sep = ""))
  N0.kluyveri.PAML.summary <- rbind(N0.kluyveri.PAML.summary, 
                               c(IGC.geo, mean(target["N0_kluyveriYDR418W", ]) - 0.215682181791, 
                                 sd(target["N0_kluyveriYDR418W", ])))
  N0.kluyveri.PAML.summary <- rbind(N0.kluyveri.PAML.summary, 
                               c(IGC.geo, -(mean(target["N0_N1", ]) - 0.01972409), 
                                 sd(target["N0_N1", ])))
  N0.kluyveri.PAML.summary <- data.frame(N0.kluyveri.PAML.summary)
  N0.kluyveri.PAML.summary <- rbind(N0.kluyveri.PAML.summary, 
                               c(IGC.geo,-(mean(target["N0_N6", ]) - 0.01972409), 
                                 sd(target["N0_N6", ])))
  N0.kluyveri.PAML.summary <- data.frame(N0.kluyveri.PAML.summary)
}
N0.kluyveri.PAML.summary <- data.frame(N0.kluyveri.PAML.summary)
N0.kluyveri.PAML.summary <- cbind(N0.kluyveri.PAML.summary, matrix(c("outgroup", "N0_N1", "N0_N6"), length(IGC.geo.list) * 3, 1))
colnames(N0.kluyveri.PAML.summary) <- c("IGC.geo", "over_estimation", "sd_over_estimation", "branch")
N0.kluyveri.PAML.summary <- data.frame(N0.kluyveri.PAML.summary)


p.full <- ggplot(N0.kluyveri.PAML.summary, aes(x = IGC.geo, y= over_estimation, colour = branch, group = branch)) +
    geom_errorbar(aes(ymin = over_estimation -  qnorm(0.975) * sd_over_estimation, ymax = over_estimation + qnorm(0.975) * sd_over_estimation)) + 
    #    geom_line() + 
    geom_point(shape = 21, fill = "white") + 
    xlab("Mean tract length") + 
    ylab("Over / - under branch length estimation") + 
    #xlim(0.0, 0.25) + #max(target[, "True_blen"], target[, "estimated_blen"] + target[, "sd_of_blen"]) * 1.1) + 
    #ylim(0.0, 0.25) + #max(target[, "True_blen"], target[, "estimated_blen"] + target[, "sd_of_blen"]) * 1.1) +
    ggtitle(paste("Over / -under estimation from PAML", sep = "")) +
    fulltheme
ggsave(paste(results.path, "OverEstimationPAML.pdf", sep = ""))
ggsave(paste(results.path, "OverEstimationPAML.jpg", sep = ""))


t.test.overestimation.summary <- NULL
for (IGC.geo in IGC.geo.list){
  target <- get(paste("geo_", toString(IGC.geo), ".0_PAML_maxlnL_summary", sep = ""))
  t.test.paralog.1 <- t.test(target["N0_kluyveriYDR418W", ] - 0.215682181791, -(target["N0_N1", ] - 0.01972409), paired = TRUE)
  t.test.paralog.2 <- t.test(target["N0_kluyveriYDR418W", ] - 0.215682181791, -(target["N0_N6", ] - 0.01972409), paired = TRUE)
  t.test.overestimation.summary <- rbind(t.test.overestimation.summary, 
                               c(IGC.geo, t.test.paralog.1$statistic[[1]], 
                                 t.test.paralog.2$statistic[[1]],
                                 t.test.paralog.1$p.value, t.test.paralog.2$p.value))
  
}
colnames(t.test.overestimation.summary) <- c("IGC.geo",  "t of outgroup vs. N0_N1", "t of outgroup vs. N0_N6",
                              "p-value of outgroup vs. N0_N1", "p-value of outgroup vs. N0_N6")

```
